---
title: example usage of amsplot.r
author: july (jpjagt)
---

first, we load the file so that we can use the functions inside.

```{r}
# assuming the file is in the same directory.
source("./amsplot.r")
```

next, we read a dataset file, and perform some basic pre-processing to convert
the categorical ranking columns to numerical columns, using the
categorical_ranking2number function provided by amsplot.r.

```{r}
# the location of the housing dataset.
df_housing_fpath <- './datasets/amsterdam_housing.csv'

# read the file.
df_housing <- read.csv(df_housing_fpath)

# satisfaction, condition, maintenance, neighborhood, development, homefeel
# are all categorical ranking columns, so convert those.
df_housing$hou_satisfaction <- categorical_ranking2number(df_housing$hou_satisfaction)
df_housing$hou_condition <- categorical_ranking2number(df_housing$hou_condition)
df_housing$hou_maintenance <- categorical_ranking2number(df_housing$hou_maintenance)
df_housing$hou_neighborhood <- categorical_ranking2number(df_housing$hou_neighborhood)
df_housing$hou_development <- categorical_ranking2number(df_housing$hou_development)
df_housing$hou_homefeel <- categorical_ranking2number(df_housing$hou_homefeel)
```

next, we have to select a single year (because we can only display one value
per neighborhood). let's make a filtered dataframe which only contains values from 2019:
```{r}
df_housing_2019 <- df_housing[df_housing$year == 2019,]
```

okay, we're ready to plot! let's first display a plot directly. for instance,
the self-reported satisfaction:

```{r}
print(amsplot_col(df_housing_2019, "hou_satisfaction"))
```

cool! what about the number of small houses (< 40m2)?

well, of course, the total number of small houses in a neighborhood depends on
the total number of houses in that neighborhood, so we can't really compare it
directly.  therefore, let's plot the average number of small houses by
normalizing the total number of small houses by the total number of houses in
each neighborhood.

```{r}
print(amsplot_col(df_housing_2019, "hou_40", normalize_by="hou_total"))
```

if we want to save a plot, we refrain from printing it, and instead use the
function ggsave:

```{r}
print(amsplot_col(df_housing_2019, "hou_satisfaction"))
ggsave("hou_satisfaction_in_2019.png")
```

that should save a PNG in the current (working) directory.
